Cores from kicks: constraints for observations of gravitational recoil in elliptical galaxies

### 2024-05-08 papers/paper-corekick/paper-corekick.tex
small edits

\subsection{Mass density profiles in three dimensions}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{density_3d}
    \caption{
        Three dimensional stellar mass density, with the line colour corresponding to the kick velocity of the SMBH. 
        All simulations with $\vk \geq 0.5v_\mathrm{esc}=900\,\kmps$ share the same colour, as there is little variation in the profiles for these kick velocities. 
        A clear dichotomy is revealed in the mass profiles, whereby high kick velocoties induce a core, and low kick velocities do not, with a sudden transition at \drafting{$\vk=600\,\kmps$}.
        }
    \label{fig:density3d}
\end{figure}
Before considering the projected, 2D stellar mass density profiles of the merger remnants, we investigate the 3D stellar mass distribution.
This alleviates the influence of projection effects on inferring if a core has formed during the BH binary merger process, and gives insight into the dependence of how matter is spatially distributed as a function of the recoiling SMBH kick velocity.
For the $\vk=0\,\kmps$ case, we determine the SMBH influence radius to be $r_\mathrm{infl}\simeq 1.5\,\mathrm{kpc}$, and is insensitive to the time of measurement.
In \autoref{fig:density3d}, we observe a clear dichotomy in the 3D spatial distribution of stellar mass, with mergers that had an SMBH with a kick velocity $\vk<540\,\kmps$ having more mass in the central $3\,\mathrm{kpc}$ ($2r_\mathrm{infl}$) than those mergers with $\vk>540\,\kmps$. 
All merger remnants have a consistent mass density distribution at $r>3\,\mathrm{kpc}$. 
Additionally, for kick velocities $\vk<540\,\kmps$, we observe a gradient in decreasing central stellar density with increasing kick velocity.
We check the time dependence of this dichotomy by constructing the 3D stellar mass distributions at additional times $t=0.056\,\mathrm{Gyr}$, $t=0.094\,\mathrm{Gyr}$, and $t=0.12\,\mathrm{Gyr}$, and find the dichotomy persists, with the same kick velocity split.

@@

To explain the presence of the bifurcation in the IFU maps, we must investigate how the stellar density is behaving in the wake of the recoiling SMBH.
We take two representative recoil velocities, $\vk=0\,\kmps$ and $\vk=1500\,\kmps$, and limit our investigation to stellar particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$ (approximately matching the aperture of the mock IFU maps), recalling that the SMBH is ejected along the positive $x$ axis.
We create velocity vector fields of the stellar particle motion in the $x$-$y$ plane \autoref{fig:wake}.
For the $\vk=0\,\kmps$ case, there is an excess of stellar motion parallel to the $x$-axis, with some orthogonal motion present at small radii. 
This is contrasted to the $\vk=1500\,\kmps$ case, where there is no clear alignment of the stellar velocity flow with either the $x$- or $y$- axes.
We interpret this feature as stellar particles diffusing into the potential wake carved by the recoiling SMBH in the $\vk=1500\,\kmps$ case, where the diffusion is neither parallel nor orthogonal to the SMBH motion, but rather at a variety of angles to it.

@@

\section{Specific IFU maps}
\drafting{Here we include both paralle and orthogonal projections of the IFU maps for three representative kick velocities: $\vk=0\,\kmps$, $\vk=600\,\kmps$, and $\vk=1800\,\kmps$.}

@@

\section{Stellar velocity fields}
We create vector fields of stellar particles for two representative merger remnants, namely $\vk=0\,\kmps$ and $\vk=1500\,\kmps$.
The velocity fields are created for particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$, as described in \autoref{sec:discussion}, and are shown in \drafting{figure}.

>>>>

\subsection{Mass density profiles in three dimensions}
\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{density_3d}
    \caption{
        Three dimensional stellar mass density, with the line colour corresponding to the kick velocity of the SMBH. 
        All simulations with $\vk \geq 0.5v_\mathrm{esc}=900\,\kmps$ share the same colour, as there is little variation in the profiles for these kick velocities. 
        A clear dichotomy is revealed in the mass profiles, whereby high kick velocities induce a core, and low kick velocities do not, with a sudden transition at $\vk=540\,\kmps$.
        }
    \label{fig:density3d}
\end{figure}
Before considering the projected, 2D stellar mass density profiles of the merger remnants, we investigate the 3D stellar mass distribution.
This alleviates the influence of projection effects on inferring if a core has formed during the BH binary merger process, and gives insight into the dependence of how matter is spatially distributed as a function of the recoiling SMBH kick velocity.
For the $\vk=0\,\kmps$ case, we determine the SMBH influence radius to be $r_\mathrm{infl}\simeq 1.5\,\mathrm{kpc}$, and is insensitive to the time of measurement.
In \autoref{fig:density3d}, we observe a clear dichotomy in the 3D spatial distribution of stellar mass, with mergers that had an SMBH with a kick velocity $\vk<540\,\kmps$ having more mass in the central $3\,\mathrm{kpc}$ ($2r_\mathrm{infl}$) than those mergers with $\vk>540\,\kmps$. 
All merger remnants have a consistent mass density distribution at $r>3\,\mathrm{kpc}$. 
Additionally, for kick velocities $\vk<540\,\kmps$, we observe a gradient in decreasing central stellar density with increasing kick velocity.
We check the time dependence of this dichotomy by constructing the 3D stellar mass distributions at additional times $t=0.056\,\mathrm{Gyr}$, $t=0.094\,\mathrm{Gyr}$, and $t=0.12\,\mathrm{Gyr}$, and find the dichotomy persists, with the same kick velocity split.

@@

To explain the presence of the bifurcation in the IFU maps, we must investigate how the stellar density is behaving in the wake of the recoiling SMBH.
We take two representative recoil velocities, $\vk=0\,\kmps$ and $\vk=1500\,\kmps$, and limit our investigation to stellar particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$ (the aperture of the mock IFU maps in \autoref{sec:app_ifu} extended to approximately $5\,\mathrm{kpc}$), recalling that the SMBH is ejected along the positive $x$ axis.
We create velocity vector fields of the stellar particle motion in the $x$-$y$ plane (see \autoref{fig:wake}).
For the $\vk=0\,\kmps$ case, there is an excess of stellar motion parallel to the $x$-axis, with some orthogonal motion present at small radii. 
This is contrasted to the $\vk=1500\,\kmps$ case, where there is no clear alignment of the stellar velocity flow with either the $x$- or $y$- axes.
We interpret this feature as stellar particles diffusing into the potential wake carved by the recoiling SMBH in the $\vk=1500\,\kmps$ case, where the diffusion is neither parallel nor orthogonal to the SMBH motion, but rather at a variety of angles to it.

@@

\section{Specific IFU maps}\label{sec:app_ifu}
\drafting{Here we include both paralle and orthogonal projections of the IFU maps for three representative kick velocities: $\vk=0\,\kmps$, $\vk=600\,\kmps$, and $\vk=1800\,\kmps$.}

@@

\section{Stellar velocity fields}
We create vector fields of stellar particles for two representative merger remnants, namely $\vk=0\,\kmps$ and $\vk=1500\,\kmps$.
The velocity fields are created for particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$, as described in \autoref{sec:discussion}, and are shown in \autoref{fig:wake}.
